---
title: "Exercise"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library (tidyverse)
library (broom)
#install.packages('lmtest')
library(lmtest)
library(nlme)

```

# Spatial and temporal structure

1. Simulate a 100-observation autocorrelated timeseries with arima.sim, with a first order autoregressive coefficient of 0.5. Also make a time vector of 1:100

2. Plot the data

```{r}
simulation <- arima.sim (n = 100, list(ar = 0.5))
simulation
plot(simulation)
```

3. Regress the timeseries against time with an OLS model. Does the model appear to be statistically significant?

How to: x: time series; y: the arima.sim model
Answer: The model does not appear to be statistically significant (p = 0.1247)

4a. Plot the model diagnostics
4b. Include an autocorrelation function and partial autocorrelation function of the residuals

```{r first, echo = TRUE}
set.seed(42)
dat <- tibble(
  x = 1:100,
  y = arima.sim (n = 100, list(ar = 0.5)))

mod <- lm(y ~ x, data = dat)
anova(mod)
summary(mod)

par(mfrow = c(2,2))
plot(mod)

acf(resid(mod))
pacf(resid(mod))
```

5. Use the Durbin-Watson test to test the residuals for autocorrelation

Result: DW = 0.97573: positive autocorrelation

```{r}
dwtest(mod)
```


6. Fit a gls with an appropriate correlation structure. Is this a better model? How have the p-value and effect size changed?

Answer: p-value increased to 0.6199; effect size changed slightly: from 0.006465 to 0.00220386

a. GLS with corAR1

```{r}
year <- time(simulation)
fit.gls <- gls(simulation ~ year, correlation = corAR1())
summary(fit.gls)
```

b. GLS without including correlation 

```{r}
year <- time(simulation)
fit.gls2 <- gls(simulation ~ year)
summary(fit.gls2)
```

c.Running an anova of both to compare 

See that the AIC of the gls with correlation included is lower = better model
```{r}
anova(fit.gls, fit.gls2)
```

7. Repeat the above 1000 times and find how autocorrelation affects the distribution of the p-value and the effect size.

```{r}

#Repeat all steps with time series 1000; simulation 2

simulation2 <- arima.sim (n = 1000, list(ar = 0.5))
simulation2
plot(simulation2)

#dat2, mod2

set.seed(42)
dat2 <- tibble(
  x = 1:1000,
  y = arima.sim (n = 1000, list(ar = 0.5)))

mod2 <- lm(y ~ x, data = dat2)
anova(mod2)
summary(mod2)

par(mfrow = c(2,2))
plot(mod2)

acf(resid(mod2))
pacf(resid(mod2))

dwtest(mod2)

#Gives DM: 1.0191; positive autocorrelation 

#GLS with correlation; gls3
#GLS without correlation; gls4

year <- time(simulation2)
fit.gls3 <- gls(simulation2 ~ year, correlation = corAR1())
summary(fit.gls3)

year <- time(simulation2)
fit.gls4 <- gls(simulation2 ~ year)
summary(fit.gls4)

anova (fit.gls3, fit.gls4)
```


Real data
The built-in dataset LakeHuron has annual lake level data from 1885 to 1972 Load the data with the command data(LakeHuron)

Plot the data.

Regress the LakeHuron lake level against year using a linear model. Is there a significant trend?

Plot the autocorrelation and partial autocorrelation functions for the residuals from this regression. Interpret them.

Fit an autoregressive models to the residuals. Compare the results with your interpretation of the PACF plot.

Fit a gls model using a corAR1 correlation structure. Test if the correlation structure is necessary. Is the trend significant?

Fit a gls model using a corARMA correlation structure with two AR terms. Is this model an improvement?